Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance in zooplankton

### 2022-05-06 text/Manuscript.Rmd
manuscript_test

---
title: "Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance in zooplankton"
author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Gáspár Jékely"
date: "*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 4QD, UK*"
output:
  pdf_document: 
      latex_engine: xelatex
      template: style.tex
  html_document: 
    theme: lumen
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
mainfont: Arial
documentclass: article
classoption:
  - onecolumn
linestretch: 1.2
fontsize: 10pt
bibliography: references.bib
csl: elife.csl

---

@@

@@

![Figure 1. Legend goes here](../figures/Figure1.png){width=80%}

>>>>

---
title: "Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance in zooplankton"
author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Gáspár Jékely"
date: "*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 4QD, UK*"
output:
  pdf_document: 
      latex_engine: xelatex
      template: style.tex
  html_document: 
    theme: lumen
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
mainfont: Arial
documentclass: article
classoption:
  - onecolumn
linestretch: 1.2
fontsize: 10pt
bibliography: references.bib
csl: elife.csl
---

@@

**Nitric oxide synthase is expressed in neurons in the cPRC circuit** First we performed a nitric oxide synthase (NOS) homology search on Platynereis transcriptome data and identified one NOS gene. We examined the gene expression pattern of NOS using in situ hybridization chain reaction (HCR). In the 2-day-old larvae, the NOS gene was strongly expressed in four cells, one pair on the dorsal side and one pair on the ventral side (Fig. 1A). NOS cells were labelled using 12 kbp upstream of the transcription start point of the NOS gene and NOS cells were identified as four interneurons in the ciliary-type photoreceptor cells (cPRCs) circuit from Platynereis reconstruction data (Fig. 1B). We named these NOS interneurons, INNOS (INNOSdl, dr, vl, vr) (Fig. 1C). The cPRC circuit has been reported as a neural circuit involved in c-opsin1 mediated UV avoidance behaviour (Verasztó et al., 2017). INNOS were presynaptic from the cPRCs and postsynaptic to the RGW interneurons in the cPRC circuit (Fig. 1D). To determine whether NO is involved in UV avoidance behaviour, we generated mutant lines with the CRISPR/Cas-9 system and examined their behavioural responses to UV light.

**NOS mutant delay from UV avoidance** We created two mutant lines with different deletions in the third exon of the NOS gene (NOSΔ11/Δ11 and Δ23/Δ23) (Fig. 2A). Compared to wild-type larvae, homozygous NOS mutant larvae have a defective UV-avoidance response (Fig. 2B). Whereas wild-type larvae showed downward swimming behaviour within 30 seconds in response to lateral UV light (UV-avoidance), the NOS mutant larvae showed a delayed onset of this behaviour and a reduced downward swimming speed (Fig. 2B). These results are similar to previously reported responses of UV-avoidance behaviour in c-opsin1 mutants (Verasztó et al., 2017), suggesting that NO is involved in UV-avoidance behaviour via the cPRC circuit. We therefore searched for target molecules for NO.

**GCs with NIT (nitrite/nitrate sensing) domains (NIT-GCs) are expressed in cPRC** Generally, NO can act via cyclic guanosine monophosphate (cGMP) as a second messenger. In this signalling pathway, NO binds to the heme group of soluble guanylate cyclase (sGC), a member of the guanylate cyclase family with a CYC domain (PF00211), to increase cGMP. However, no NO targets were found in cells that highly express sGC within the cPRC circuit. Recently, Moroz and others reported that GC with NIT (nitrite/nitrate sensing) domains (PF08376) (NIT-GC), identified as a sensor for nitrate and nitrite, the oxides of NO, is relatively widely conserved in metazoans (Moroz et al., 2020). To search for NIT-GCs in Platynereis, genes with the CYC domains were searched using the transcriptome and genome sequences of 45 metazoan species and choanoflagellate. By sequence-similarity searches, multiple sequence alignments and cluster analyses, a cluster of membrane-bound (guanylate cyclase-coupled receptor) type was found and a phylogenetic tree was constructed with soluble type (sGC) as an outgroup. As in the phylogenetic analysis of Moroz et al, we found a group of GCs with NIT domains that included many metazoans such as placozoa, cnidaria, ecdysozoa, echinoderms and lophotrochozoa, and in Platynereis 15 NIT-GCs were found (Fig. 3A). To find the candidate NIT-GCs involved in the cPRC circuit among the 15 NIT-GCs, we used previously published spatially mapped single-cell transcriptome data (Achim et al., 2015; Williams et al., 2017) and Two NIT-GCs were found to be highly expressed in the cPRC, and we analysed their expression using in situ HCR as NIT-GC1 and NIT-GC2. As a result, strong signals for both were observed specifically in the vicinity of the cPRCs in 2-day-old larvae (Fig. 3B). Both c-opsin1 and MLD pNP (pedal peptide neuropeptide precursor 2), already identified as cPRC markers, and NIT-GC1 and NIT-GC2 were found to be co-expressed. They were co-expressed c-opsin1 and MLD pNP (pedal peptide neuropeptide precursor 2), already identified as cPRC markers (Fig. 3C) (Arendt et al., 2004; Williams et al., 2017). To analyse localisations at the protein level, immunostaining was performed using polyclonal antibodies against NIT-GC1 and NIT-GC2. Interestingly, NIT-GC1 was found to localise to the neuronal axon and NIT-GC2 to the ciliary region of the cPRC (Fig. 3D and E). Furthermore, co-immunostaining with NOS antibodies revealed that NOS and NIT-GC1 were localised in close proximity in the neurosecretory plexus (Fig. 3F). These results suggest that NIT-GC1 may function as a major target for NO.

**NIT-GC1 produces cGMP by nitric oxide** NIT-GC1 has a highly conserved cyclase domain and activation of NIT-GC1 is expected to produce cGMP. Recently, a synthetic protein called Green cGull was developed as a cGMP indicator that can be used in cultured cell systems (Matsuda et al., 2017). We co-expressed Green cGull and NIT-GC1 in cultured cells COS-7 (monkey kidney) and tested the intracellular dynamics of cGMP in the activation of NIT-GC1. First, S-nitroso-N-acetyl-DL-penicillamine (SNAP; NO donor), 8-Br-cGMP (8-Bromoguanosine 3',5'-cyclic monophosphate; membrane-permeable cGMP analogue, positive control) and DMSO (negative control) were applied to COS-7 cells expressing Green cGull only. 8-Br-cGMP increased fluorescence intensity by more than sixfold, while little changes were detected for SNAP and DMSO (Fig. 4A). Nest, the two kind of genes, Green cGull and NIT-GC1, were co-expressed using the 2A self-cleaving peptide and the same solutions were applied. As a result, a clear increase in cGMP fluorescence intensity was detected in SNAP compared to DMSO or expression of Green cGull only, although not as high as the increase in 8-Br-cGMP (Fig. 4B). Finally, expression of a mutant form of NIT-GC1 with deletion of the NIT domain significantly suppressed the increase in cGMP (Note, we need re-experiment!, Fig. 4C). These results indicate that NIT-GC1 is activated by NO to produce cGMP. They also show that the NIT domain is required for the activation of NIT-GC1.

**NO induces depolarisation of the cPRCs via NIT-GC1** To determine the function of NO in the cPRC circuit, the intracellular dynamics of neurons in the cPRC circuit were investigated, focusing on NIT-GC1. First, we investigated the changes in NO during UV stimulation of the cPRC using DAF-FM, a fluorescent probe that can detect NO, in order to determine whether NO is produced by UV light. By injecting DAF-FM directly into zygotes, INNOS cells could be identified under live imaging. To visualise the location of cPRCs, the calcium indicator RGECO1a was also injected together, as cPRCs maintain a higher resting calcium levels compared to other cells. Direct UV stimulation to the cPRCs increased the fluorescence intensity of DAF-FM compared to stimulation of non-cPRC region (Fig. 5A). Next, we did calcium imaging of cells in the cPRC circuit during UV stimulation. We locally stimulated the cPRC cilia in wild-type with a 405 nm laser (UV) and detected a slight increase followed by a strong decrease in calcium levels (hyperpolarisation) (Fig. 5B). Turning off the UV after or during the decrease caused a strong increase (depolarisation) and a return to the same calcium level as before (Fig. 5B). These responses were similar to those reported previously (Verasztó et al., 2017). We performed a similar stimulation using NOS mutants, which hyperpolarised during the UV stimulation, but no depolarising responses were observed when UV was stopped (Fig. 5C). Similarly, in NIT-GC1 knockdown larvae using the morpholinos, no depolarisation after the hyperpolarisation response were observed (Fig. 5D). Interestingly, no hyperpolarising response during UV stimulation was found to occur in NIT-GC2 knockdown larvae. And we compared the changes of the INNOS calcium levels identified by their position in relation to cPRCs during UV stimulation in wild-type and in NOS mutants. In both cases, a depolarising response was observed at the same time as the hyperpolarising response of the cPRCs (Fig. 5E). Furthermore, to perform calcium imaging of INRGWs in UV stimulation, INRGWs were identified by immunostaining with RGWa antibodies after calcium imaging (Fig. 5F). It has been reported that the INRGW cells form many large synapses with synaptic vesicles on two serotonergic ciliomotor neurons (Ser-h1) that are known to play a role in the regulation of ciliary beating (Verasztó et al., 2017) and the activity of the INRGW cells increased the negative correlation between the activity patterns of serotonergic neurons and the MC neuron (Williams et al., 2018). The results showed that depolarisation of INNOS occurs at the same time as hyperpolarisation of cPRC in the wild type. Identified INRGW cells were found to cause depolarisation at the same time as cPRC hyperpolarisation in the wild type (Fig. 5G). It was also found that in NOS mutants where cPRC hyperpolarisation does not occur, the INRGW depolarisation response also does not occur (Fig. 5H).

@@

![Figure 1. Legend goes here](../figures/Figure1.png){width="80%"}
